---
title: "Lab 10 - Grading the professor"
author: "Anaelle Gackiere"
date: "03-27-2026"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

Here is a link to the [lab
instructions](https://datascience4psych.github.io/DataScience4Psych/lab10.html).

## Load Packages and Data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

# Part 1

## Exercise 1

The distribution of scores is negatively skewed, so students are
relatively generous. The mean score is 4.175, and the minimum score
given was 2.3 (at least not a 0!).

```{r exercise1_code}
# descriptives
summary(evals$score)

# visualization
ggplot(evals, aes(x = score)) +
  geom_bar( fill = "plum4", bins = 30) +
  theme_minimal()

```

## Exercise 2

I actually initially did geom_jitter, but doing geom_point is
problematic for discrete variables, and will result in a strange looking
scatterplot like this one below (see description at exercise 3). The
issue with geom_point here is that we are not (EXPLAIN).

```{r exercise2_code}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(title = "Profesor Evaluation Score vs Average Beauty",
       x = "Average Professor Beauty Rating",
       y = "Evaluation Score") +
  theme_classic() 
```

## Exercise 3

```{r exercise3-code}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(title = "Profesor Evaluation Score vs Average Beauty",
       x = "Average Professor Beauty Rating",
       y = "Evaluation Score") +
  theme_classic() 

```

We have a few outliers of low scores with low attractiveness, but
overall most of the points are high regardless of beauty level. After
closer inspection, there might be a somewhat positive trend between
beauty and score, but it's unclear to what extent and whether or not
it's linear.

# Part 2

### Exercise 4

The fitted model is: **y = 3.88 + 0.0666 × bty_avg**
For every one-point increase in average beauty rating, the model
predicts evaluation score increases by about 0.07 points. For exampple,
if a professor goes from a beauty rating of 4 to a 5, their predicted
evaluation score goes up by 0.07.

```{r regression-model}
# model
m_bty <- lm(score ~ bty_avg, data = evals)

# view output
m_bty

# effect check
glance(m_bty)

```

### Exercise 5

```{r regression-plot}
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_jitter(width = 0.1, height = 0.1, color = "black", alpha = 0.5) +
  geom_smooth(method = "lm", color = "orange", se = FALSE) +
  labs(
    title = "Evaluation Score vs. Beauty Rating",
    subtitle = "Regression Line in Orange",
    x = "Average Professor Beauty Rating",
    y = "Evaluation Score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

### Exercise 6

As mentioned before, for every one-point increase in average beauty
rating, the model predicts evaluation score increases by about 0.07
points; thus, as beauty ratings increase, so do evaluation scores. The
intercept is not meaningful here since there is no true 0. This is a
statistically detectable but practically trivial effect, since beauty
explains only 3.5% of the variation in scores (R² = 0.035), meaning the
vast majority of what drives evaluation scores has nothing to do with
looks. Right now, it seems that beauty is a minor piece of the story.
The issue with the shading here is it can make a trivial relationship
look like a big deal just because it takes up visual space.

# Part 3

### Exercise 7

The reference level is "Female"; R can't put "female" and "male" into a
math equation directly, so it turns one into 0 and the other into 1.
Female = 0 (the baseline), male = 1. The two coefficients are the
Intercept = 4.09, which is the predicted score for female professors
(the reference level) and gendermale = 0.14. Male professors score 0.14
points higher than female professors on average

Predicted mean evaluation score for each gender:

Female: 4.09 + 0.14(0) = 4.09

Male: 4.09 + 0.14(1) = 4.23

```{r exercise7-code}

m_gen <- lm(score ~ gender, data = evals)
tidy(m_gen)

```

### Exercise 8

Now let’s do one more categorical predictor: rank

Actually, let’s do three slightly different versions of this model to
see how changing the reference level or regrouping categories affects
the output.

Create a new variable from rank called rank_relevel where "tenure track"
is the baseline level. Create another new variable called
tenure_eligible that labels "teaching" faculty as "no" and labels
"tenure track" and "tenured" faculty as "yes".

Fit three different linear models to explore how rank affects evaluation
scores. Each of these models will use a different version of the rank
variable to predict average professor evaluation score

```{r exercise8-code}
# tenure track, tenure eligible




```

### Exercise 9

Based on the regression outputs, interpret how teaching faculty and
tenured faculty differ from that baseline. (Hint you should interpret
the slopes and intercepts for all three models in context of the data.)

### Exercise 10

Report the\
R 2 . Does rank explain much, or not really?

Part 4: Multiple Linear Regression \### Is the “beauty effect” still
there once we account for gender?

Fit these two models:

m_bty: score \~ bty_avg

m_bty_gen: score \~ bty_avg + gender

```{r}




```

Then answer:

What changes in the beauty slope when gender is added?

For two professors with the same beauty rating, does gender still shift
the predicted score?

Compare the adjusted\
R 2 values. Is gender actually helping much, or is beauty doing most of
the work already?

Finally, swap gender out and add rank instead.

Fit m_bty_rank: score \~ bty_avg + rank and then interpret one rank
coefficient and the beauty slope in context.

Part 5: The Search for the Best Model Going forward, only consider the
following variables as potential predictors: rank, ethnicity, gender,
language, age, cls_perc_eval, cls_did_eval, cls_students, cls_level,
cls_profs, cls_credits, bty_avg.

Which variable, on its own, would you expect to be the worst predictor
of evaluation scores? Why? Hint: Think about which variable would you
expect to not have any association with the professor’s score.

Check your suspicions from the previous exercise. Include the model
output for that variable in your response.

Suppose you wanted to fit a full model with the variables listed above.
If you are already going to include cls_perc_eval and cls_students,
which variable should you not include as an additional predictor? Why?

Fit a full model with all predictors listed above (except for the one
you decided to exclude) in the previous question.

```{r}



```

Using backward-selection with adjusted R-squared as the selection
criterion, determine the best model. You do not need to show all steps
in your answer, just the output for the final model. Also, write out the
linear model for predicting score based on the final model you settle
on.

```{r}


```

Interpret the slopes of one numerical and one categorical predictor
based on your final model.

Based on your final model, describe the characteristics of a professor
and course at University of Texas at Austin that would be associated
with a high evaluation score.

Would you be comfortable generalizing your conclusions to apply to
professors generally (at any university)? Why or why not?

I would be hesitant to generalize my conclusions to professors at any
university, as students and professors are nested in the university, so
I would want to run an MLM to test this across multiple universities,
and then we may be able to generalize more confidently to the general
university population.
