---
title: "Lab 10 - Grading the professor"
author: "Anaelle Gackiere"
date: "03-27-2026"
output: github_document
---

Here is a link to the [lab instructions](https://datascience4psych.github.io/DataScience4Psych/lab10.html).

## Load Packages and Data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

Part 1

## Exercise 1

The distribution of scores is negatively skewed, so students are relatively generous. The mean score is 4.175, and the minimum score given was 2.3 (at least not a 0!).

```{r exercise1_code}
# descriptives
summary(evals$score)

# visualization
ggplot(evals, aes(x = score)) +
  geom_bar( fill = "plum4", bins = 30) +
  theme_minimal()

```

## Exercise 2
I actually initially did geom_jitter, but doing geom_point is problematic for discrete variables, and will result in a strange looking scatterplot like this one below (see description at exercise 3). The issue with geom_point here is that we are not (EXPLAIN). 

```{r exercise2_code}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(title = "Profesor Evaluation Score vs Average Beauty",
       x = "Average Professor Beauty Rating",
       y = "Evaluation Score") +
  theme_classic() 
```

## Exercise 3
```{r exercise3-code}
ggplot(data = evals, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(title = "Profesor Evaluation Score vs Average Beauty",
       x = "Average Professor Beauty Rating",
       y = "Evaluation Score") +
  theme_classic() 

```
We have a few outliers of low scores with low attractiveness, but overall most of the points are high regardless of beauty level. After closer inspection, there might be a somewhat positvie trend between beauty and score, but it's unclear to what extent and whether or not it's linear. 

Part 2

### Exercise 4

Equation for the fitted model: score = 3.88 + 0.0666 x bty_avg. A one-point increase in beauty predicts a 0.07 increase for evaluation scores. The effect is quite small: R-squared = 0.035. That is, professors rated as more attractive receive slightly higher evaluation scores.
```{r regression-model}
# model
m_bty <- lm(score ~ bty_avg, data = evals)

# view output
m_bty

# effect check
glance(m_bty)

```

Replot your existing visualization, this time add a regression line in orange. Turn off the default shading around the line. (By default, the plot includes shading around the line.)

### Exercise 5

```{r regression-plot}
ggplot(evals, aes(x = bty_avg, y = score)) +
  geom_jitter(width = 0.1, height = 0.1, color = "black", alpha = 0.5) +
  geom_smooth(method = "lm", color = "orange", se = FALSE) +
  labs(
    title = "Evaluation Score vs. Beauty Rating",
    subtitle = "Regression Line in Orange",
    x = "Average Professor Beauty Rating",
    y = "Evaluation Score"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```

### Exercise 6
Then, stepping back, interpret what this model is really saying:

How much do evaluation scores change with beauty ratings (slope)?
What does the intercept represent here? Is it meaningful in this context, or just a mathematical artifact? Explain your reasoning.
What is the  
R
2
  value of this model? Interpret it in context. What does it tell you about whether beauty is a “big deal” or a pretty minor piece of the story?
That shading you turned off represents uncertainty around the line—why might it be easy to read too much into it at this stage?

Part 3

### Exercise 7

Take a look at the model output. What’s the reference level? What do the coefficients tell you about how evaluation scores differ between male and female professors? Write the predicted mean evaluation score for each gender.

```{r exercise7-code}

m_gen <- lm(score ~ gender, data = evals)
tidy(m_gen)

```

### Exercise 8

Now let’s do one more categorical predictor: rank

Actually, let’s do three slightly different versions of this model to see how changing the reference level or regrouping categories affects the output.

Create a new variable from rank called rank_relevel where "tenure track" is the baseline level. Create another new variable called tenure_eligible that labels "teaching" faculty as "no" and labels "tenure track" and "tenured" faculty as "yes".

Fit three different linear models to explore how rank affects evaluation scores. Each of these models will use a different version of the rank variable to predict average professor evaluation score


### Exercise 9
Based on the regression outputs, interpret how teaching faculty and tenured faculty differ from that baseline. (Hint you should interpret the slopes and intercepts for all three models in context of the data.)

### Exercise 10 
Report the  
R
2
 . Does rank explain much, or not really?


Part 4: Multiple Linear Regression
### 
Is the “beauty effect” still there once we account for gender?

Fit these two models:

m_bty: score ~ bty_avg

m_bty_gen: score ~ bty_avg + gender

Then answer:

What changes in the beauty slope when gender is added?

For two professors with the same beauty rating, does gender still shift the predicted score?

Compare the adjusted  
R
2
  values. Is gender actually helping much, or is beauty doing most of the work already?

Finally, swap gender out and add rank instead.

Fit m_bty_rank: score ~ bty_avg + rank and then interpret one rank coefficient and the beauty slope in context.

Part 5: The Search for the Best Model
Going forward, only consider the following variables as potential predictors: rank, ethnicity, gender, language, age, cls_perc_eval, cls_did_eval, cls_students, cls_level, cls_profs, cls_credits, bty_avg.

Which variable, on its own, would you expect to be the worst predictor of evaluation scores? Why? Hint: Think about which variable would you expect to not have any association with the professor’s score.

Check your suspicions from the previous exercise. Include the model output for that variable in your response.

Suppose you wanted to fit a full model with the variables listed above. If you are already going to include cls_perc_eval and cls_students, which variable should you not include as an additional predictor? Why?

Fit a full model with all predictors listed above (except for the one you decided to exclude) in the previous question.

Using backward-selection with adjusted R-squared as the selection criterion, determine the best model. You do not need to show all steps in your answer, just the output for the final model. Also, write out the linear model for predicting score based on the final model you settle on.

Interpret the slopes of one numerical and one categorical predictor based on your final model.

Based on your final model, describe the characteristics of a professor and course at University of Texas at Austin that would be associated with a high evaluation score.

Would you be comfortable generalizing your conclusions to apply to professors generally (at any university)? Why or why not?

